---
title: "Understanding the elements in output"
author: Yuan-Ling Liaw and Waldir Leoncio
header-includes:
    - \usepackage{setspace}\onehalfspacing
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Understanding the elements in output}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include = FALSE, warning = FALSE}
library(knitr)
options(width = 90, tidy = TRUE, warning = FALSE, message = FALSE)
opts_chunk$set(comment = "", warning = FALSE, message = FALSE,
               echo = TRUE, tidy = TRUE)
```

```{r load}
library(lsasim)
```

```{r packageVersion}
packageVersion("lsasim") 
```

```{r equation, eval=FALSE}
questionnaire_gen(n_obs, cat_prop = NULL, n_vars = NULL, n_X = NULL, n_W = NULL, 
                  cor_matrix = NULL, cov_matrix = NULL, 
                  c_mean = NULL, c_sd = NULL, 
                  theta = FALSE, family = NULL, 
                  full_output = FALSE, verbose = TRUE)
```

If the logical argument `full_output` is `TRUE`, output will be a list containing the questionnaire data as well as several objects that might be of interest for further analysis of the data. 

---

We generate one continuous and two ordinal covariates. We specify the covariance matrix between the numeric and ordinal variables. The data is generated from a multivariate normal distribution. And we set the logical argument `full_output = TRUE`. 

```{r, include = FALSE}
set.seed(1234)
(props <- list(1, c(.25, 1), c(.2, .8, 1)))
(yw_cov <- matrix(c(1, .5, .5, .5, 1, .8, .5, .8, 1), nrow = 3))
bg <- questionnaire_gen(n_obs = 10, cat_prop = props, cov_matrix = yw_cov, theta = TRUE, 
                  family = "gaussian", full_output = TRUE)
names(bg)
```

The output is a list containing the following elements: `r names(bg)`.

```{r, eval=FALSE}
?questionnaire_gen
```

```{r ex 1a}
set.seed(1234)
(props <- list(1, c(.25, 1), c(.2, .8, 1)))
(yw_cov <- matrix(c(1, .5, .5, .5, 1, .8, .5, .8, 1), nrow = 3))
```

```{r}
questionnaire_gen(n_obs = 10, cat_prop = props, cov_matrix = yw_cov, theta = TRUE, 
                  family = "gaussian", full_output = TRUE)
```

